On the specific recursive structure of iterated parametricity

### 2025-11-16 paper/paper.tex
paper: improve last equation
paper: abbreviate with depscoht

\newcommandx{\deps}[2][1,2]{\depsp{}[][][][#1][#2][][]}
\newcommandx{\fulldeps}[1][1]{\depsp{full}[][][][#1][][][]}
\newcommandx{\depscohs}[2][1,2]{\depsp{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohs}[1][1]{\depsp{fullcohs}[][#1][][][]}
\newcommandx{\depscoht}[2][1,2]{\depsp{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscoht}[1][1]{\depsp{fullcoht}[][][][#1][][][]}
\newcommandx{\depstype}[2][1,2]{\depstypep{}[][][][#1][#2][][]}
\newcommandx{\fulldepstype}[2][1,2]{\depstypep{full}[][][][#1][][][]}
\newcommandx{\depscohstype}[2][1,2]{\depstypep{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohstype}[1][1]{\depstypep{fullcohs}[][][][#1][][][]}
\newcommandx{\depscohttype}[2][1,2]{\depstypep{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscohttype}[1][1]{\depstypep{fullcoht}[][][][#1][][][]}

@@

This is enough to define $\cohf[n][p]$, using $\UIP$ to solve the dependency in $\coht[n][p]$.

@@

@@

@@

                \coht[n][0,...,p-1]:\cohttype[n][0,...,p-1]
              \end{array}\right) : \cohftype[n+1][0,...,p]}                                                                \\
            \end{array}\right)              & \defeq & (\unitpoint,\lambda q.\lambda r. \lambda\epsilon. \lambda\omega.\lambda \_.\refl) \\
              \coht[n][0,...,p] \\
                  \coht[n+1][0,...,p-1] \\
                \end{array}\right),                                        \\
                  \coht[n+1][0,...,p-1] \\
                \end{array}\right)(d), \cohl[n][p][q][r][\epsilon][\omega]
                  \coht[n][0,...,p] \\
                \end{array}\right)(d)(l))                                    \\
            \end{array}\right)      : \U}                                                                \\

@@

                  \coht[n][0,...,p-1]

@@

                  \coht[n][0,...,p-1]

@@

                  \coht[n][0,...,p-1]

@@

              \cohp[n][0,...,p]:\cohptype[n][0,...,p] \\
              \coht[n][0,...,p]:\cohttype[n][0,...,p] \\
            \end{array}\right)                                 :

@@

      \end{array}} \\

@@

              \cohf[n][0,...,p] \\

@@

This is time for a new abbreviation:

\begin{align*}
  \depscohttype[n][0,..,p-1] \defeq
  \left(\begin{array}{lll}
            \fulldepscohs[n]    & : & \fulldepscohstype[n]                      \\
            \cohp[n][0,...p-1]  & : & \cohptype[n][0,...,p-1](\fulldepscohs[n]) \\
            \coht[n][0,...,p-1] & : & \cohttype[n][0,...,p-1]
            \left(\begin{array}{l}
                \fulldepscohs[n]    \\
                \cohp[n][0,...,p-1] \\
              \end{array}\right)                                           \\
          \end{array}\right)
\end{align*}

It is also convenient to build $\depscohs[n+1][0,...,p]$ from $\depscoht[n][0,...,p]$ at the previous level:

\begin{align*}
  \overline{\depscohs}(\depscoht[n][0,...,p]:\depscohttype[n][0,...,p]) & :      & \depscohstype[n+1][0,...,p] \\
  \overline{\depscohs}(\depscoht[n][0,...,p])                           & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}(\depscoht.\fulldepscohs.\depscohs[n][0,...,p-1]) \\
            \restrp[n+1][p](\depscoht.\fulldepscohs[n][0,...,n])              \\
            \cohf[n+1][p](\depscoht[n][0,...,p-1])                            \\
          \end{array}\right)                                      \\
\end{align*}

@@

\begin{align*}
  \depscohs[n][p]\begin{cases}
                   \deps[n][p]\begin{cases}
                 \framep[n][0,...,n-1]   & : \frametype[n][0,...,n-1]                           \\
                 \painting[n][0,...,n-1] & : \paintingtype[n][0,...,n-1](\framep[n][0,...,n-1]) \\
                 \restrf[n][0,...,n-1]   & : \restrftype[n][0,...,n-1]
                 \left(\begin{array}{l}
                \framep[n][0,...,n-1]   \\
                \painting[n][0,...,n-1] \\
              \end{array}\right)
               \end{cases} \\
                   \restrp[n][0,...,n-1](E:\framep[n+1][n]
                   \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
            \end{array}\right)\rightarrow \U) & : \restrptype[n][0,...,n-1]
                   \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
              E                       \\
            \end{array}\right)                                                             \\
                   \cohf[n][0,...,n-1](E:\framep[n+1][n]
                   \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
            \end{array}\right)\rightarrow \U) & : \cohftype[n][0,...,n-1]
                   \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
              E                       \\
              \restrp[n][0,...,n-1]   \\
            \end{array}\right)
                 \end{cases} \\
  \cohp[n][0,...,n-1]
  \left(\begin{array}{l}
            E:\framep[n+1][n]
            \left(\begin{array}{l}
                \framep[n][0,...,n-1]   \\
                \painting[n][0,...,n-1] \\
                \restrf[n][0,...,n-1]   \\
              \end{array}\right)\rightarrow \U \\
            E':\framep[n+2][n+1](\overline{\deps}
            \left(\begin{array}{l}
                \framep[n][0,...,n-1]   \\
                \painting[n][0,...,n-1] \\
                \restrf[n][0,...,n-1]   \\
                E                       \\
                \restrp[n][0,...,n-1]   \\
                \cohf[n][0,...,n-1]     \\
              \end{array}\right))\rightarrow \U
          \end{array}\right)
  : \cohptype[n][0,...,n-1]
  \left(\begin{array}{l}
            \framep[n][0,...,n-1]   \\
            \painting[n][0,...,n-1] \\
            \restrf[n][0,...,n-1]   \\
            E                       \\
            \restrp[n][0,...,n-1]   \\
            \cohf[n][0,...,n-1]     \\
            E'                      \\
          \end{array}\right)                                                                           \\
\end{align*}

>>>>

\newcommandx{\deps}[2][1,2]{\depsp{}[][][][#1][#2][][]}
\newcommandx{\fulldeps}[1][1]{\depsp{full}[][][][#1][][][]}
\newcommandx{\depscohs}[2][1,2]{\depsp{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohs}[1][1]{\depsp{fullcohs}[][][][#1][][][]}
\newcommandx{\depscoht}[2][1,2]{\depsp{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscoht}[1][1]{\depsp{fullcoht}[][][][#1][][][]}
\newcommandx{\depstype}[2][1,2]{\depstypep{}[][][][#1][#2][][]}
\newcommandx{\fulldepstype}[2][1,2]{\depstypep{full}[][][][#1][][][]}
\newcommandx{\depscohstype}[2][1,2]{\depstypep{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohstype}[1][1]{\depstypep{fullcohs}[][][][#1][][][]}
\newcommandx{\depscohttype}[2][1,2]{\depstypep{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscohttype}[1][1]{\depstypep{fullcoht}[][][][#1][][][]}

@@

Thanks to $\cohptype[n][0,...,p-1]$, we can mutually define $\cohf[n+1][0,...,p]$ and $\cohttype[n][0,...,p-1]$ from $\fulldepscohs[n][0,...,p-1]$, abbreviating this and the dependency in a $\coht[n][0,...,p-1]$ as $\depscohttype[n][0,...,p-1]$, and using also abbreviations in passing for $\cohl$ and $\cohltype$. It is also convenient to build $\depscohs[n+1][0,...,p]$ from $\depscoht[n][0,...,p-1]$, so that we can now build $\framep[n+3][p]$. Thus, we also include this abbreviation in the definition:

\begin{align*}

@@

    \depscohttype[n][0,..,p-1]                                       &                                                   & \defeq                                                                            &
    \left(\begin{array}{lll}
              \fulldepscohs[n]    & : & \fulldepscohstype[n]                      \\
              \cohp[n][0,...p-1]  & : & \cohptype[n][0,...,p-1](\fulldepscohs[n]) \\
              \coht[n][0,...,p-1] & : & \cohttype[n][0,...,p-1]
    \overline{\depscohs}                                             & (\depscoht[n][0,...,p]:\depscohttype[n][0,...,p]) & :                                                                                 & \depscohstype[n+1][0,...,p] \\
    \overline{\depscohs}                                             & \depscoht[n][0,...,p]                             & \defeq                                                                            &
              \overline{\deps}(\depscoht.\fulldepscohs.\depscohs[n][0,...,p-1]) \\
              \restrp[n+1][p](\depscoht.\fulldepscohs[n][0,...,n])              \\
              \cohf[n+1][p](\depscoht[n][0,...,p-1])                            \\

@@

    (\fulldepscoht[n]:\fulldepscohttype[n])                          & :                                                 & \cohftype[n+1][0,...,p]                                                                                         \\
    (\unitpoint)                                                     & \defeq                                            & (\unitpoint,\lambda q.\lambda r. \lambda\epsilon. \lambda\omega.\lambda \_.\refl)                               \\
    (\fulldepscoht[n])                                               & \defeq                                            &
              (\fulldepscoht[n]),                                                \\
              (\fulldepscoht[n])(d), \cohl[n][p][q][r][\epsilon][\omega]
              (
              \fulldepscoht[n])(d)(l))                                           \\

@@

            \end{array}\right) & :                                                 & \U                                                                                                                                                      \\

@@

              (\fulldepscoht[n])

@@

              (\fulldepscoht[n+1])

@@

              (\fulldepscoht[n])                                                    \\

@@

    (\fulldepscoht[n]:\fulldepscohttype[n])                          & :                                                 &

@@

    \end{array}                                                                                                     \\

@@

    (\fulldepscoht[n])                                               & \defeq                                            &

@@

\end{align*}

@@

\begin{align*}
  \rotatebox[origin=c]{90}{\depscohs[n][0,...,n-1]}\begin{cases}
                                                     \rotatebox[origin=c]{90}{\deps[n][0,...,n-1]}\begin{cases}
                                                   \framep[n][0,...,n-1]   : \frametype[n][0,...,n-1]                           \\
                                                   \painting[n][0,...,n-1] : \paintingtype[n][0,...,n-1](\framep[n][0,...,n-1]) \\
                                                   \restrf[n][0,...,n-1]   : \restrftype[n][0,...,n-1]
                                                   \left(\begin{array}{l}
                \framep[n][0,...,n-1]   \\
                \painting[n][0,...,n-1] \\
              \end{array}\right)
                                                 \end{cases} \\
                                                     E:\framep[n+1][n]
                                                     (\deps[n][0,...,n-1])\rightarrow \U                                                                                       \\
                                                     \restrp[n][0,...,n-1](E) : \restrptype[n][0,...,n-1]
                                                     \left(\begin{array}{l}
              \deps[n][0,...,n-1] \\
              E                   \\
            \end{array}\right)                                                                                                 \\
                                                     \cohf[n][0,...,n-1](E)   : \cohftype[n][0,...,n-1]
                                                     \left(\begin{array}{l}
              \deps[n][0,...,n-1]                 \\
              E:\framep[n+1][n]
              (\deps[n][0,...,n-1])\rightarrow \U \\
              \restrp[n][0,...,n-1]               \\
            \end{array}\right)
                                                   \end{cases} \\
  E': \framep[n+2][n+1](\overline{\deps}
  (\depscohs[n][0,...,n-1])\rightarrow \U                                                                                                                                                  \\
  \cohp[n][0,...,n-1]
  \left(\begin{array}{l}
            E  \\
            E' \\
          \end{array}\right)
  : \cohptype[n][0,...,n-1]
  \left(\begin{array}{l}
            \depscohs[n][0,...,n-1] \\
            E'                      \\
          \end{array}\right)                                                                                                                                                            \\
\end{align*}
